Evaluation of probabilistic photometric redshift estimation approaches for \textsc{LSST}

### 2019-10-14 paper/tex/analysis.tex
clarified description of CDF metrics

We evaluate a number of quantitative metrics derived from the visually interpretable QQ plot and PIT histogram, built on the Kolmogorov-Smirnov (KS) statistic
\begin{equation}
  \label{eq:ks}
  \mathrm{KS} \equiv \max_{z} \left( \left| \mathrm{CDF}[\hat{f}, z] - \mathrm{CDF}[\tilde{f}, z] \right| \right),
\end{equation}
interpretable as the maximum difference between the CDFs of an approximating univariate distribution $\hat{f}(z)$ and a reference distribution $\tilde{f}(z)$, in this case $U(0,1)$.
% $CDF_{\rm phot}(z)$ and $CDF_{\rm spec}(z)$, the CDFs of the photo-$z$ and spectroscopic redshift respectively:
%The KS test quantifies whether the two redshift distributions are drawn from a similar distribution, independent of binning.
% The KS test quantifies the similarity between two distributions, independent of binning.
% A lower $N_{\rm KS}$ value corresponds to more similar distributions.
%We calculate two KS test statistics, one to test the distribution of $p(z)$ values, and one to test the stacked estimator $\hat{n}(z)$ of the redshift distribution.
%The $p(z)$ test compares the CDF of the PIT values described in the previous section to that of a uniform distribution.
%The $n(z)$ test compares the CDF of the distribution formed by stacking the $p(z)$ for every object in our test set against the CDF of the corresponding true redshifts.
%In this case a lower $N_{\rm KS}$ value means that the photo-$z$ distribution fits the spectroscopic redshift distribution well. $F_{\rm phot}(z)$ is calculated by stacking the PDFs of every single object in the sample, producing a smooth cumulative $n(z)$ to be compared with $F_{\rm spec}(z)$.
We also consider two variants of the KS statistic.
A cousin of the KS statistic, the Cramer-von Mises (CvM) statistic
\begin{equation}
\label{eq:cvm}
  \mathrm{CvM}^{2} \equiv \int_{-\infty}^{+\infty} \big(\mathrm{CDF}[\hat{f}, z] - \mathrm{CDF}[\tilde{f}, z]\big)^2 \mathrm{d}\mathrm{CDF}[\tilde{f}, z]
\end{equation}
is the mean-squared difference between the CDFs of an approximate and true PDF.
The Anderson-Darling (AD) statistic
\begin{equation} \label{eq:ad}
  \mathrm{AD}^2 \equiv N_{tot}\int_{-\infty}^{+\infty} \frac{\big(\mathrm{CDF}[\hat{f}, z] - \mathrm{CDF}[\tilde{f}, z]\big)^2} {\mathrm{CDF}[\tilde{f}, z] (1 -\mathrm{CDF}[\tilde{f}, z]} \mathrm{d}\mathrm{CDF}[\tilde{f}, z]
\end{equation}
is a weighted mean-squared difference featuring enhanced sensitivity to discrepancies in the tails of the distribution.
%The KS statistic and the CvM and AD statistics related to it are all used to evaluate the performance of both $\hat{n}(z)$ and the QQ and PIT of the CDFs of the ensemble of photo-$z$ interim posteriors.
In anticipation of a substantial fraction of galaxies having PIT of 0 or 1, a consequence of catastrophic outliers, we evaluate the AD statistic with modified bounds of integration $(0.01, 0.99)$ to exclude those extremes in the name of numerical stability.

>>>>

We evaluate a number of quantitative metrics derived from the visually interpretable QQ plot and PIT histogram, built on the Kolmogorov-Smirnov (KS) statistic
\begin{equation}
  \label{eq:ks}
  \mathrm{KS} \equiv \max_{z} \left( \left| \mathrm{CDF}[\hat{f}, z] - \mathrm{CDF}[\tilde{f}, z] \right| \right),
\end{equation}
interpretable as the maximum difference between the CDFs of the empirical distribution of PIT values for the test sample $\hat{f}(z)$ and a reference distribution $\tilde{f}(z)$, in this case $U(0,1)$, for the ideal distribution of PIT values.
% $CDF_{\rm phot}(z)$ and $CDF_{\rm spec}(z)$, the CDFs of the photo-$z$ and spectroscopic redshift respectively:
%The KS test quantifies whether the two redshift distributions are drawn from a similar distribution, independent of binning.
% The KS test quantifies the similarity between two distributions, independent of binning.
% A lower $N_{\rm KS}$ value corresponds to more similar distributions.
%We calculate two KS test statistics, one to test the distribution of $p(z)$ values, and one to test the stacked estimator $\hat{n}(z)$ of the redshift distribution.
%The $p(z)$ test compares the CDF of the PIT values described in the previous section to that of a uniform distribution.
%The $n(z)$ test compares the CDF of the distribution formed by stacking the $p(z)$ for every object in our test set against the CDF of the corresponding true redshifts.
%In this case a lower $N_{\rm KS}$ value means that the photo-$z$ distribution fits the spectroscopic redshift distribution well. $F_{\rm phot}(z)$ is calculated by stacking the PDFs of every single object in the sample, producing a smooth cumulative $n(z)$ to be compared with $F_{\rm spec}(z)$.
We also consider two variants of the KS statistic.
A cousin of the KS statistic, the Cramer-von Mises (CvM) statistic
\begin{equation}
\label{eq:cvm}
  \mathrm{CvM}^{2} \equiv \int_{-\infty}^{+\infty} \big(\mathrm{CDF}[\hat{f}, z] - \mathrm{CDF}[\tilde{f}, z]\big)^2 \mathrm{d}\mathrm{CDF}[\tilde{f}, z]
\end{equation}
is the mean-squared difference between the CDFs of an approximate and true PDF.
The Anderson-Darling (AD) statistic
\begin{equation} \label{eq:ad}
  \mathrm{AD}^2 \equiv N_{tot}\int_{-\infty}^{+\infty} \frac{\big(\mathrm{CDF}[\hat{f}, z] - \mathrm{CDF}[\tilde{f}, z]\big)^2} {\mathrm{CDF}[\tilde{f}, z] (1 -\mathrm{CDF}[\tilde{f}, z]} \mathrm{d}\mathrm{CDF}[\tilde{f}, z]
\end{equation}
is a weighted mean-squared difference featuring enhanced sensitivity to discrepancies in the tails of the distribution.
%The KS statistic and the CvM and AD statistics related to it are all used to evaluate the performance of both $\hat{n}(z)$ and the QQ and PIT of the CDFs of the ensemble of photo-$z$ interim posteriors.
In anticipation of a substantial fraction of galaxies having PIT of 0 or 1, a consequence of catastrophic outliers, we evaluate the AD statistic with modified bounds of integration $(0.01, 0.99)$ to exclude those extremes in the name of numerical stability.
